---
title: 'IMT 573: Problem Set 3 - Working With Data II'
author: "Wenyue Yin"
date: 'Due: Tuesday, October 22, 2019'
output:
  html_document:
    df_print: paged
---

<!-- This syntax can be used to add comments that are ignored during knitting process. -->

##### Collaborators: <!-- BE SURE TO LIST ALL COLLABORATORS HERE! -->

##### Instructions:

Before beginning this assignment, please ensure you have access to R and RStudio; this can be on your own personal computer or on the IMT 573 R Studio Server. 

1. Download the `problemset3.Rmd` file from Canvas or save a copy to your local directory on RStudio Server. Open `problemset3.Rmd` in RStudio and supply your solutions to the assignment by editing `problemset3.Rmd`. 

2. Replace the "Insert Your Name Here" text in the `author:` field with your own full name. Any collaborators must be listed on the top of your assignment. 

3. Be sure to include well-documented (e.g. commented) code chucks, figures, and clearly written text chunk explanations as necessary. Any figures should be clearly labeled and appropriately referenced within the text. Be sure that each visualization adds value to your written explanation; avoid redundancy -- you do not need four different visualizations of the same pattern.

4.  Collaboration on problem sets is fun and useful, and we encourage it, but each student must turn in an individual write-up in their own words as well as code/work that is their own.  Regardless of whether you work with others, what you turn in must be your own work; this includes code and interpretation of results. The names of all collaborators must be listed on each assignment. Do not copy-and-paste from other students' responses or code.

5. All materials and resources that you use (with the exception of lecture slides) must be appropriately referenced within your assignment.  

6. Remember partial credit will be awarded for each question for which a serious attempt at finding an answer has been shown. Students are \emph{strongly} encouraged to attempt each question and to document their reasoning process even if they cannot find the correct answer. If you would like to include R code to show this process, but it does not run without errors, you can do so with the `eval=FALSE` option as follows:

```{r example chunk with a bug, eval=FALSE}
a + b # these object dont' exist 
# if you run this on its own it with give an error
```

7. When you have completed the assignment and have **checked** that your code both runs in the Console and knits correctly when you click `Knit PDF`, rename the knitted PDF file to `ps3_YourLastName_YourFirstName.pdf`, and submit the PDF file on Canvas.

##### Setup: 

In this problem set you will need, at minimum, the following R packages.

```{r}
# Load standard libraries
library('dplyr')
library('censusr')
library('stringr')
```

#### Problem 1: Joining census data to police reports

In this problem set, we will be joining disparate sets of data - namely: Seattle police crime data, information on Seattle police beats, and education attainment from the US Census. Our ultimate goal is to build a dataset where we can examine questions around crimes in Seattle and the educational attainment of people living in the areas in which the crime occurred.

As a general rule, be sure to keep copies of the original dataset(s) as you work through cleaning (remember data provenance).

##### (a) Importing and Inspecting Crime Data

Load the Seattle crime data (crime_data.csv). You can find more information on the data here: (https://data.seattle.gov/Public-Safety/Crime-Data/4fs7-3vj5). This dataset is constantly refreshed online so we will be using the csv file for consistency. We will henceforth call this dataset the "Crime Dataset." Perform a basic inspection of the Crime Dataset and discuss what you find.

```{r}
raw_data = read.csv("Crime_Data.csv")
```

##### (b) Looking at Years That Crimes Were Committed

Let's start by looking at the years in which crimes were committed. What is the earliest year in the dataset? Are there any distinct trends with the annual number of crimes committed in the dataset?

```{r}
data_year <- group_by(raw_data, year = substr(raw_data$Occurred.Date,7,10))
year <- summarise(data_year, count= n())
year

#So the earliest year in the dataset is 1908. The annual number is quite stable these years, but there was a sudden rise in 2008, probably because the change in recording reports. And there was a rise during 2013 to 2014. As the data is till May 2019, it's reasonable that the annual number of 2019 is less than 20 thousand.
```

Let's subset the data to only include crimes that were committed after 2011 (remember good practices of data provenance!). Going forward, we will use this data subset.

```{r}
data <- filter(data_year, year >= 2011)
```

##### (c) Looking at Frequency of Beats

How frequently are the beats in the Crime Dataset listed? Are there any anomolies with how frequently some of the beats are listed? Are there missing beats?
```{r}
data_beat <- group_by(data, Name = Beat)
data_beat <- summarise(data_beat, count= n())
data_beat

# The frequences of the beats are quite similar in the same section, but there are few missing beats that have different beat names in data.
```


##### (d) Importing Police Beat Data and Filtering on Frequency

Load the data on Seattle police beats (police_beat_and_precinct_centerpoints.csv). You can find additional information on the data here: (https://data.seattle.gov/Land-Base/Police-Beat-and-Precinct-Centerpoints/4khs-fz35) and more information on what a police beat is here: https://en.wikipedia.org/wiki/Beat_(police). We will henceforth call this dataset the "Beats Dataset."

```{r}
# add your R code here
raw_beat_dataset = read.csv("Police_Beat_and_Precinct_Centerpoints.csv")

```

Does the Crime Dataset include police beats that are not present in the Beats Dataset? If so, how many and with what frequency do they occur? Would you say that these comprise a large number of the observations in the Crime Dataset or are they rather infrequent? Do you think removing them would drastically alter the scope of the Crime Dataset?

```{r}

beat_ljoin <- left_join(data_beat, raw_beat_dataset, by = "Name")
beat_missingbeats <- filter(beat_ljoin, is.na(Latitude) == TRUE)
select(beat_missingbeats, Name, count)

#There are 7 missing beats that are not in the beats dataset. However, as these beats have a quite small number of the observation of crime, I think removing them would not alter the scope.
```


Let's remove all instances in the Crime Dataset that have beats which occur fewer than 10 times across the Crime Dataset. Also remove any observations with missing beats. After only keeping years of interest and filtering based on frequency of the beat, how many observations do we now have in the Crime Dataset?

```{r}
beat10 <- filter(beat_ljoin, count <= 10)
data_cleaned <- filter(data, Beat != "CTY",Beat != "DET", Beat != "INV", Beat != "K",Beat != "N",Beat != "S",Beat != "SS",Beat != "W",Beat != "WS",Beat != "")
data_cleaned
# Actually I was thinking about joining the datasets and then filter the observations, but the dataset is a bit large and I'm not sure if that will take a lot of space and time, so I just filter the 10 beats by hand.

# Answer : There are 389002 observations after 2011.
```

##### (e) Importing and Inspecting Police Beat Data

To join the Beat Dataset to census data, we must have census tract information. 

First, let's remove the beats in the Beats Dataset that are not listed in the (cleaned) Crime Dataset.

Then, let's use the censusr package to extract the 15-digit census tract for each police beat using the corresponding latitude and longitude. Do this using each of the police beats listed in the Beats Dataset. Do not use a for-loop for this but instead rely on R functions (e.g. the 'apply' family of functions). Add a column to the Beat Dataset that contains the 15-digit census tract for the each beat. (HINT: you may find censusr's call_geolocator_latlon function useful)

```{r}
#Remove the beats that are not listed in the crime dataset, which means to remove the beats that have observation <10.
beat <- left_join(raw_beat_dataset, beat10, by = "Name")
beat <- filter(beat, is.na(beat$count) == TRUE)
beat <- select(beat, Name, Location = Location.1.x, Latitude = Latitude.x, Longitude =Longitude.x)

myfun <- function(Lat,Lon){return(call_geolocator_latlon(Lat,Lon))}

beat_censustract <- mutate(beat, census_tract = mapply(myfun, beat$Latitude,beat$Longitude))

beat_censustract
```

We will eventually join the Beats Dataset to the Crime Dataset. We could have joined the two and then found the census tracts for each beat. Would there have been a particular advantage/disadvantage to doing this join first and then finding census tracts? If so, what is it? (NOTE: you do not need to write any code to answer this)

# Maybe the speed will be much slower.


##### (f) Extracting FIPS Codes

Once we have the 15-digit census codes, we will break down the code based on information of interest. You can find more information on what these 15 digits represent here: https://transition.fcc.gov/form477/Geo/more_about_census_blocks.pdf.

First, create a column that contains the state code for each beat in the Beats Dataset. Then create a column that contains the county code for each beat. Find the FIPS codes for WA State and King County (the county of Seattle) online. Are the extracted state and county codes what you would expect them to be? Why or why not?

```{r}
beats <- mutate(beat_censustract, state_code = substr(beat_censustract$census_tract,1,2),county_code = substr(beat_censustract$census_tract,3,5))
beats

# The FIPS codes for WA is 53, for King County is 033. So the result is what I have expected, as the dataset contains Seattle crime data.
```



##### (g) Extracting 11-digit Codes

The census data uses an 11-digit code that consists of the state, county, and tract code. It does not include the block code. To join the census data to the Beats Dataset, we must have this code for each of the beats. Extract the 11-digit code for each of the beats in the Beats Dataset. The 11 digits consist of the 2 state digits, 3 county digits, and 6 tract digits. Add a column with the 11-digit code for each beat.

```{r}
beats <- mutate(beats, digital_code_11 = substr(beat_censustract$census_tract,1,11))
beats
```

##### (h) Extracting 11-digit Codes From Census

Now, we will examine census data (census_edu_data.csv). The data includes counts of education attainment across different census tracts. Note how this data is in a 'wide' format and how it can be converted to a 'long' format. For now, we will work with it as is.

The census data contains a "GEO.id" column. Among other things, this variable encodes the 11-digit code that we had extracted above for each of the police beats. Specifically, when we look at the characters after the characters "US" for values of GEO.id, we see encodings for state, county, and tract, which should align with the beats we had above. Extract the 11-digit code from the GEO.id column. Add a column to the census data with the 11-digit code for each census observation.

```{r}
raw_edu <- read.csv("census_edu_data.csv")
edu_11_digital_code <- mutate(raw_edu, digital_code_11 = substr(raw_edu$GEO.id,10,21))
edu_11_digital_code

#There's already a column called geo.id2, which is exactly the same as the column I added on the table.
```

##### (i) Join Datasets

Join the census data with the Beat Dataset using the 11-digit codes as keys. Be sure that you do not lose any of the police beats when doing this join (i.e. your output dataframe should have the same number of rows as the cleaned Beats Dataset - use the correct join). Are there any police beats that do not have any associated census data? If so, how many?

```{r}
beat_edu <- left_join(beats, edu_11_digital_code, by = "digital_code_11" )

beat_edu %>%
  filter(total>0) %>%
  summarise(count = n())

#All police beats have the associated census data, as the number of the beats that the total number of education > 0 is equal to the number of rows of the joint table.
```



Then, join the Crime Dataset to our joined beat/census data. We can do this using the police beat name. Again, be sure you do not lose any observations from the Crime Dataset. What is the final dimensions of the joined dataset?

```{r}
crime_beat_census <- mutate(beat_edu, Beat = Name)
crime_beat_census <- left_join(data_cleaned, crime_beat_census, by = "Beat")
crime_beat_census <- select(crime_beat_census, -Name)
crime_beat_census


```


Once everything is joined, save the final dataset for future use.